\section{Results}
\label{sec:results}

\subsection{Word boundary prediction}
Table x shows the precision scores for word boundary prediction and word onset detection. When looking at the models that take the activation patterns of the third hidden layer as input, precision is only marginally better than a default prediction of 'no word boundary', which would be correct in .72 of cases. % insert exact number 
However, when we look at the lower layers, precision is considerably better. [Comparison to performance of n-gram models here]
These results indicates that sub-word linguistic structure is mainly handled in the lower layers.

\begin{table}[]
	\centering
	\caption{Table x. Precision of word boundary prediction and detection}
	\begin{tabular}{lll}
		& Word ends      & Onsets         \\
		Layer 1 & 0.8864   & 0.9173 \\
		Layer 2 & 0.8376  & 0.8589 \\
		Layer 3 & 0.7672 & 0.7789
	\end{tabular}
\end{table}

\subsection{Word Similarity} % you may want to say semantic, although that is somewhat questionable, and relatedness rather than similarity...
Table y shows Spearmans \rho between human similarity ratings from the MEN dataset and the cosine similarity at the last timestep for all layers. The results in the first column are based on the full dataset, and the second column only includes word pairs for which both words appeared at least 10 times in the training portion of MSCOCO.
The pattern in the similarity as judged by the different layers shows that the higher the layer, the more semantic information is incorporated: there is virtually no information in the lowest hidden layer, and it is strongest in the third layer. However, even at layer 3 and for well-known words, Spearmans \rho is .31. Clearly, although the correlation is there, it is only weak.  % compare to some other results? How bad really is this? How come?

\begin{table}[]
	\centering
	\caption{Table y. Spearmans \rho between cosine similarity and human similarity judgements} % bad title
	\begin{tabular}{lll}
		& All words      & Frequent words only         \\
		Layer 1 & 0.0712 & 0.0671 \\
		Layer 2 & 0.2077 & 0.2254 \\
		Layer 3 & 0.2787 & 0.3072 
	\end{tabular}
\end{table}

% do you normally report p here?
%Layer   well-known words        all words
%1       SpearmanrResult(correlation=0.067060499075493871, pvalue=0.00076678081271022865)        SpearmanrResult(correlation=0.071211128496393786, %pvalue=9.4723361217079462e-05)
%2       SpearmanrResult(correlation=0.22538353323450167, pvalue=2.5609637334957168e-30) SpearmanrResult(correlation=0.20774408372704065, pvalue=1.3261150819357359e-30)
%3       SpearmanrResult(correlation=0.30718765478550603, pvalue=4.4310485063154974e-56) SpearmanrResult(correlation=0.27868574652167188, pvalue=1.2141988368305471e-54)

% report on lev and stuff... LOTS TO DO
